---
output:
  rmarkdown: github_document
  html_document: default
  pdf_document: default
---

# Data Structures and Subsetting

In this tutorial we are trying to demonstrate basic operations with data structures and subsetting in R software. This file is part of my project for Statistical Computing with Prof. Harner. 
#### Explain your code where appropriate.

1. Create vectors of each of the different primitive types. i.e., integer, double, logical, and character. Create matrices by attaching `dim` attributes to those vectors. Look up the help for `dimnames` and attach `dimnames` to these resulting matrices.

```{r}
# Put your R code here.
int_v <- c(10L,1L,2L)
double_v <- c(10.1,10.2,10.3)
char_v <- c("a","b","c")
log_v <- c(TRUE,FALSE,F)

# Adding dim attributes - First Method
dim(int_v) <- c(1,3)
dim(double_v) <- c(1,3)
dim(char_v) <- c(1,3)
dim(log_v) <- c(1,3)

#second way of making matrices
matrix(int_v, c(1,3))
matrix(double_v, c(1,3))
matrix(char_v, c(1,3))
matrix(log_v, c(1,3))

#dimnames
dimnames(int_v) <- list(c("r1"),c("Col1","Col2","Col3"))
int_v
dimnames(double_v) <- list(c("r1"),c("Col1","Col2","Col3"))
double_v
dimnames(char_v) <- list(c("r1"),c("Col1","Col2","Col3"))
char_v
dimnames(log_v) <- list(c("r1"),c("Col1","Col2","Col3"))
log_v
```

2. Create a list of length 4 and then add a `dim` attribute to it. What happens?

```{r}
# Put your R code here.
l0=list(c(1,2,3,4),"ali",c(1L,2L),c(T,F))
length(l0)
dim(l0) <- c(2,2)
l0
typeof(l0)
# The list is showing in a matrix form of 2*2. It's type is list.
```


3. Look up the help page for `data.frame` and use the example code to create a small data frame containing numeric variables and factors.

```{r}
# Put your R code here.
month = c("March","April","January","November","January", 
 "September","October","September","November","August",
 "January","November","November","February","May","August",
 "July","December","August","August","September","November",
 "February","April")
fac_month = factor(month)
#fac_month
table(fac_month)
#levels(facmons)
L3 <- LETTERS[1:3]
d <- data.frame(7:30, seq(1,24), fac_month, sample(L3, 24, replace = TRUE))
d
```

4. Use the `seq` function to generate a subscript vector that selects those elements of a vector that have even-numbered subscripts.

```{r}
# Put your R code here.
x <- 10:140

#First Method:
even <- which(x %% 2 == 0)
x[even]


#Second Method:

for (i in seq(1,length(x)))
(if ( x[i] %% 2 == 0) print(x[i]) )
```

5. Verify that vectors can have duplicated names and that if a subscript matches a duplicated name, only the first value is returned. What happens with `x[NA]`?

```{r}
# Put your R code here.
x <- c(a1=1,a2=2,a1=3,a3=4)
names(x)
x["a1"]
#Only the first value is given back

x[NA]
#NA are given back for both names and values by this command
```


6. Use logical subscripts to extract the even-numbered elements of the `letters` vector.

```{r}
# Put your R code here.

cc <- letters[1:10]
cc
#First Method
for(i in 1:length(cc))
  ( if (i%% 2==0) print(cc[i])
)
  
#Second Method
remove(vector)
for(i in 1:length(cc)) #c[1]:lenght(c)
if(i %% 2 == 0)
vector <- c(vector, cc[i])
vector
```

7. Let `x` be a vector of length 10 generated by `1:10` and suppose it has a dimension attribute so that it is a matrix with 2 columns and 5 rows. What is the matrix location of the 7th element of `x`? That is, which row and column is it in? Alternatively, which element of `x` is in the second row, first column?

```{r}
# Put your R code here.
x=1:10
dim(x) <- c(2,5)
x
x[7]
# the 7th element is at the 1st row and 4th column and its value is 7. 

#========
#nrow(x)
#ncol(x)
x[2,1]
#Element 6 is in the 2nd row 1st column

#====
# if I want to find which element of matrix is equal to 11 I can use this command
which(x  == 6)
```



8. What does `as.matrix()` do when applied to a data frame with columns of different types?

```{r}
# Put your R code here.
L3 <- LETTERS[1:3]
d1 <- data.frame(1, 1:10, sample(L3, 10, replace = TRUE))
d1
as.matrix(d1)

#The elements of it will coerced to the most flexible type. According  to the help file, types from least to most flexible are: logical, integer, double, and character. So here d1's elements will be changed to char.

```

9. Fix each of the following common data frame subsetting errors:

```{r}
#mtcars[mtcars$cyl = 4, ]
#mtcars[-1:4, ]
#mtcars[mtcars$cyl <= 5]
#mtcars[mtcars$cyl == 4 | 6, ]

#Corrected Answer:

mtcars[mtcars$cyl == 4, ]
mtcars[-c(1:4), ]
mtcars[mtcars$cyl <= 5, ]
mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ]
```

10. Consider the linear model: `mod <- lm(mpg ~ wt, data = mtcars)`. Describe the data structure of `mod`, including its componets. Extract the coefficients, residuals, and the residual degrees of freedom. Extract the R squared from the model summary, i.e., from `summary(mod)`.

```{r}
#  Put your R code here.
mod <- lm(mpg ~ wt, data = mtcars)
str(mod) # We can get information about the data structure by using this command line.After running the code it will provide all the information about the structure of different variables of mod. For example it is saying that there is a list of 12 elemnts. The first element in the list is a number of 2 and its name is coefficients. Then it shows its attributes' names in the following line. After that it is giving information about the residuals which is a consisted of 32 elements(numbers) and then their attributes' names and so on...
summary(mod)  # We can get some information about the lm by using this command line

mod$df.residual
residuals(mod) #or  mod$residual
coefficients(mod) # or mod$coefficients
summary(mod)$r.squared
```


 
 